Understanding the nanoscale structure of hexagonal phase lyotropic liquid crystal membranes

### 2017-12-18 Ben_Manuscripts/structure_paper/Draft.tex
Comments added.

@@

  %BJC : I don't know that this paragraph fits.  
  Improved RO membranes can make desalination more sustainable. Only 0.5 \% of
  the world's water is fresh. Desalination is an important source of potable
  water, necessary in order to keep up with demand generated by a rapidly growing
  global population. Although RO is currently the cheapest form of desalination,
  the cost must be further decreased in order to meet demand. A typical seawater
  desalination process requires feed pressures in the range of 800 to 1000 psi
  \cite{fritzmann_state---art_2007}. By designing better membranes for
  desalination, we can achieve higher water fluxes and reduce feed pressure
  requirements which will decrease the cost of fresh water production.
  \cite{humplik_nanostructured_2011}.

@@

  Nanostructured membranes can bypass many of the performance issues which
  plague traditional NF and RO membranes. One can accomplish targeted separations
  with high selectivity by tuning shape, size and functionality of the molecular
  building blocks which form these materials. As a result, solute rejecting pores
  can have their sizes tuned uniformly, resulting in strict size cut-offs.
  Entirely different mechanisms may govern transport in a given nanostructured
  material which can inspire novel separation techniques.

@@

  Self assembling lyotropic liquid crystals (LLCs) are a suitable candidate for
  aqueous separation applications. LLCs share the characteristic ability of
  nanostructured membrane materials to create highly ordered structures with the
  added benefits of low cost and synthetic techniques feasible for large scale
  production.  \cite{feng_scalable_2014} LLC systems created by the monomer 
  Na-GA3C11 (Fig.~\ref{fig:monomer}) have been extensively studied experimentally.
  \cite{smith_ordered_1997,zhou_supported_2005,resel_h2-phase_2000,feng_scalable_2014,feng_thin_2016}
  Neat liquid crystal monomer forms the thermotropic, Col\textsubscript{h}
  phase (Fig.\ref{fig:assembly}). The presence of ca. 10 wt \% water results in the
  H\textsubscript{II} phase. In both cases, monomers assemble into mesophases
  made of hexagonally packed, uniform size, cylinders with hydrophilic head
  groups oriented inward towards the cylinder center.  The hydrophilic region can
  act as a pore for aqueous separations.  One can envision tailoring the pore
  region for specific separations by changing the monomer chemsitry.

@@

  %BJC: TODO: Scalable vector graphic for monomer structure, make bigger. Convert .svg to .eps
  %           Thicken slabs
  %           Make pore region on slab (blue) bigger
  %           Bump up atom sizes in atomistic render
  \begin{figure}
	\centering
	\begin{subfigure}{.3\textwidth}
		\centering
		\includegraphics[width=\textwidth]{NaGA3C11.png}
		\caption{}~\label{fig:monomer}
	\end{subfigure}
	\begin{subfigure}{.3\textwidth}
		\centering
		\includegraphics[width=\textwidth]{monomer_twocolor.png}
		\caption{}~\label{fig:atomistic_monomer}
	\end{subfigure}
	\begin{subfigure}{0.3\linewidth}
		\centering
		\includegraphics[width=\textwidth]{wedge.png}
		\caption{}~\label{fig:wedge}
	\end{subfigure}
		\begin{subfigure}{0.4\linewidth}
		\centering
		\includegraphics[width=\textwidth]{layer_blender.png}
		\caption{}~\label{fig:wedge_layer}
	\end{subfigure}
	\begin{subfigure}{0.4\linewidth}
		\centering
		\includegraphics[width=\textwidth]{hexagonal_packing.png}
		\caption{}~\label{fig:hex_packing_simple}
	\end{subfigure}
	\caption{The liquid crystal monomer Na-GA3C11 (a) rendered atomistically (b)
	exhibits wedge-like character (c). Monomer wedges assemble into disks (d) with
	hydrophilic head groups (blue) facing towards the disk center. The disks
	assemble into hexagonally packed columnnar mesophases (e)}~\label{fig:assembly}
  \end{figure}

  Our current understanding of LLC systems is not rich enough to be able to
  precisely design membranes for specific separations. Over the past 20 years,
  H\textsubscript{II} phase LLC membrane studies have been limited primarily to
  Na-GA3C11 with some characterization done after minor structural modifications.
  Resel et al. varied the length of the monomer tails and the counterion used and
  observed its affect on pore spacing \cite{resel_structural_2000}.  Their study
  offers no insight into dynamics within the pore. In a later study of rejection
  performance, it was shown that membranes formed by Na-GA3C11 can not perform
  separations of solutes less than 1.2 nm in diameter because the pores are too
  large \cite{zhou_supported_2005}.  We do not yet understand how to controllably
  reduce the effective pore size or how to tune the chemical environment in the
  nanopores for effective water desalination and small organic molecule
  separations. The only source of predictive modeling for LLC systems have been
  macroscopic models which likely do not adequately describe transport at these
  length scales \cite{hatakeyama_water_2011}. It will be challenging to
  efficiently narrow down the large design space in a laboratory setting without
  a robust model.

  A molecular level understanding of LLC membrane structure, enabled by
  molecular dynamics simulations, will provide guidelines to reduce the large
  chemical space available to design monomers for creation of separation-specific
  membranes. A good molecular model should incorporate a detailed picture of the
  nanoscopic pore structure which will be crucial to understanding the role of
  monomer structure in solute transport and membrane design. Models resulting
  from molecular dynamics simulations will provide the required level of detail (Fig. \ref{fig:detail}).
  We can directly observe solute transport and suggest governing mechanisms. We
  can observe how the choice of head group may influence pore size for size
  exclusion driven separations. We can interchange counterions which may
  influence both the pore size and the strength of the Donnan potential which
  affects the degree to which the membrane can exclude charged species.

  % BJC: TODO: Make new rendering of wedges
  %            Change color scheme 
  \begin{figure}
  \centering
	\begin{subfigure}{0.45\linewidth}
		\centering
		\includegraphics[width=\textwidth]{nanopore_undetailed.jpg}
		\caption{}~\label{fig:undetailed_pore}
	\end{subfigure}
	\begin{subfigure}{0.45\linewidth}
		\centering
		\includegraphics[width=\textwidth]{detailed_pore.png}
		\caption{}~\label{fig:detailed_pore}
	\end{subfigure}
  \caption{(a) Our previous understanding of the pore structure allows us to speculate
	   about separation behavior. (b) A detailed molecular model will allow us to
	   directly observe solute transport. Here, four stacked layers of 5 monomers
           are pictured atomistically. The hydrophilic region is in red and the 
	   hydrophoic region is colored blue.}~\label{fig:detail}
  \end{figure}

@@

  The $\pi$-$\pi$ stacking interactions may be a driving force of self assembly
  in this system \cite{gazit_possible_2002}. Gas phase ab initio studies of
  benzene dimers have shown a clear energetic advantage for parallel displaced
  and T-shaped $\pi$-$\pi$ stacking conformations versus a sandwiched
  conformation ~\cite{sinnokrot_estimates_2002}. Substituted benzene rings
  exhibit an even stronger $\pi$-$\pi$ stacking attraction which favors the
  parallel displaced configuration in all cases except where the substitutions
  are extremely electron withdrawing.
  \cite{waller_hybrid_2006,ringer_effect_2006}. We will compare simulated X-ray
  diffration patterns to experiment in order to deduce which stacking 
  configurations is most likely.

@@

  There remains the possibility that there is more than one metastable state
  associated with a given LLC system. Simulating a membrane atomistically will
  require many atoms which limits the timescales acessible with MD. It is
  reasonable to expect that we will generate configurations which are kinetically
  trapped in a metastable free energy basin. We must be able to identify which
  state is produced experimentally.

@@

  While the Col\textsubscript{h} phase is described as dry, it has been
  suggested by experimentalists, in unpublished communications, that it is likely
  that small amounts of ambient water are leached into neat monomer.
  Experimentally, achieving a hexagonal phase with a completely dry system has
  proven difficult. If neat monomer is allowed to sit in ambient conditions, its
  color turns from transparent to slightly opaque and a hexagonal phase forms.
  Although we will not explore whether water is necessary for self assembly, we
  hypothesize that the hydrogen bonding network formed by the water may play a
  role in structuring the pores and holding together the hexagonal phase. We can
  use simulated X-ray diffraction patterns to see if there is any meaningful
  structural difference between a "dry" and "wet" system.

@@

  %BJC: Would it be better to describe each reflection as a numbered list (like with the questions above?)
  We used experimental small-angle X-ray scattering (SAXS) data from
  \cite{feng_thin_2016} (Fig. ~\ref{fig:SAXS}) and wide angle X-ray scattering
  (WAXS) data (Fig. ~\ref{fig:WAXS}, produced as described in
  \cite{feng_scalable_2014}) for comparison to our model. We rely primarily on the 2D WAXS data
  since it encodes all structural details down to the sub-nm scale.  There are
  five major features of interest present in the 2D experimental pattern shown in
  Figure \ref{fig:WAXS}. The first is located at $q_z$ = 1.7 \angstrom$^{-1}$,
  corresponding to a real space separation of 3.7 \angstrom~.  The reflection is
  attributed to $\pi$-$\pi$ stacking between aromatic rings in the direction
  perpendicular to the membrane plane, or z-axis \cite{feng_scalable_2014}. For
  simplicity, this reflection will be referred to as R-$\pi$. A weak intensity
  line is located at exactly half the $q_z$ value of R-$\pi$ ($q_z$ = 0.85
  \angstrom$^{-1}$), corresponding to a real space periodic spacing of 7.4
  \angstrom~. This reflection has been interpreted as 2\textsubscript{1} helical
  ordering of aromatic rings along the z axis meaning if the positions of the
  aromatic rings can be traced by a helix, then for each full turn in the helix, one
  will encounter two aromatic rings. For this reason it will be referred to as
  R-helix. A third major reflection is marked by a low intensity ring located at
  r = 1.4 \angstrom$^{-1}$. The real space separation corresponds to 4.5 \angstrom~
  which is characteristic of the average spacing between packed alkane chains.
  This reflection will be called R-alkanes. Within R-alkanes, are four spots of
  higher relative intensity which will be called R-spots. All are located
  $\approx 37$ degrees from the $q_z$ axis in their respective quadrants. In many
  liquid crystal systems this can be explained by the tilt angle of the alkane
  chains with respect to the membrane plane. The final feature corresponds to the
  spacing and symmetry of the d\textsubscript{100} plane which can be related to
  the distance between pores.  The feature, which will be called R-pores, is
  characterized by dots along $q_z$ = 0. The spacing between dots is indicative
  of the hexagonal symmetry of the packed pores. The same information at higher
  resolution is obtained using a SAXS setup.

@@

  %BJC: TODO: update colorbar
  \begin{figure}
        \centering
        \begin{subfigure}[t]{0.43\linewidth}
                \centering
                \includegraphics[width=\linewidth]{SAXS.png}
                \caption{}\label{fig:SAXS}
        \end{subfigure}
        \begin{subfigure}[t]{0.47\linewidth}
                \centering
                \includegraphics[width=\linewidth]{WAXS_annotated.png} 
                \caption{}\label{fig:WAXS}
        \end{subfigure}
	\caption{(a) The repeat spacing in the 1D small angle X-ray scattering pattern
		is characteristic of hexagonal packing. The leading peak represents the
		distance between the d\textsubscript{100} planes. Using this distance, we know
		that the distance between pore centers is 4.12 nm. (b) 2D wide angle X-ray
		scattering gives details about repeating features less than 1 nanometer
		apart. Experimentalists have justified each of the 5 major reflections
		present as follows: (I) Aromatic head group $\pi-\pi$ stack 3.7 \AA~apart.
		(II) Monomers arrange vertically in a 2\textsubscript{1} helix. (III) Alkane
 		chain tails pack 4.5 \AA~apart. (IV) Monomer tails tilt with respect to the 
 		membrane plane. (V) As derived from SAXS, pores pack hexagonally and are 
		spaced 4.12 nm apart}
	\label{fig:SWAXS}
 \end{figure}

@@

  An ensemble of characteristic, low-energy vacuum monomer configurations
  were constructed by applying a simulated annealing process to a parameterized
  monomer. Monomers were cooled from 1000K to 50K over 10 nanoseconds.  A low
  energy configuration was randomly pulled from the trajectory and charges were
  reassigned using molcharge.  Using the new charges, the monomer system was
  annealed again and a random monomer configuration was pulled from the
  trajectory to be used for full system construction (Figure~\ref{fig:python}a).

@@

  A typical simulation volume contains four pores in a monoclinic unit cell,
  the smallest unit cell that maintains hexagonal symmetry when extended
  periodically. Each pore is made of twenty stacked monomer layers with periodic
  continuity in the z direction, avoiding any edge effects and creating an
  infinite length pore ideal for studying transport. A small number of layers is
  preferred in order to reduce computational cost and to allow us to look at
  longer timescales. Ultimately, we chose to build a system with 20 monomer
  layers in each pore in order to obtain sufficient resolution when simulating
  X-ray diffraction patterns. %This point will be explained in more detail later.

\subsection{Monomer Placement}

@@

  We chose the layer spacing for the initial configuration based on R-$\pi$ and
  then allowed the system to readjust during equilibration. Each monomer was rotated
  so the plane of the aromatic head groups would be coplanar with the xy plane.
  We explore two different initial layer spacings. The first is exactly equal to
  R-$\pi$ with layers placed so aromatic rings are stacked 3.7 \angstrom~ apart
  in the z-direction. A second system is explored with an initial layer spacing
  of 5 \angstrom. A third system with an initial layer spacing of 10 \angstrom was
  briefly explored. When layers are spaced out too far, they will collapse
  on each other while simultaneously slipping in between layers of adjacent pores
  which leads to an artificially thick membrane with pores spaced closely
  together. There will be no further discussion of this system but the interested
  reader can learn more about it in the supplemental information.

  The relative interlayer orientation was chosen based on clues from
  diffraction data as well as the various known stacking modes of benzene and
  substituted benzene rings: sandwiched, parallel-displaced and T-shaped
  ~\cite{sinnokrot_estimates_2002} (\Cref{fig:sandwiched,fig:pd,fig:tshaped}).
  The T-shaped configuration was ruled out because its $\approx$ 5 \angstrom~
  equilibrium stacking distance ~\cite{sinnokrot_estimates_2002} is inconsistent
  with R-$\pi$. It is also unfeasible for the monomers to orient in the T-shaped
  conformation because of the bulky tail groups. The system's preference towards
  the sandwiched vs. parallel displaced stacking modes will be explored in some
  detail.  Both have reported stacking distances near the R-$\pi$ value of 3.7
  \angstrom. Headgroups in our sandwiched initial configuration are stacked
  directly on top of each other while stacked headgroups in the parallel
  displaced initial configuration are offset by $180/nmon$ degrees where $nmon$
  equals the number of monomers per layer.

@@

  We developed equilibration schemes to create dry and wet configurations. Both
  schemes start with an initial configuration generated according to the previous
  guidelines. To create a dry configuration, we fix monomer head groups in the
  sandwiched or parallel-displaced configuration using position restraints with a
  force constant of 1e6 KJ mol$^{-1}$ nm$^{-2}$. We run a 50 ps simulation in the
  NVT ensemble which allows the monomer tails to settle without disrupting the
  ordering of the head groups. Doing so also mitigates system dependence on
  initial monomer configuration. Every 50 ps, we reduce the force constants by
  the square root of its previous value. Once the force constant is below 10 KJ
  mol$^{-1}$ nm$^{-2}$, the restraints are released linearly until there is no more
  restraining potential. The resulting unrestrained structure is allowed to
  equilibrate for 5 ns in the NPT ensemble with pressure controlled by the
  berendsen barostat. Next, we run long NPT equilibration simulations for at
  least 400 ns using the Parrinello-Rahman barostat with a time consant of 10 ps.

  In order to create a wet system, we solvated an initial configuration with
  water using gmx solvate. All water molecules placed outside the pore region are
  removed. Waters inside the pore region are randomly removed until the desired
  concentration of water in the pores is reached. The remainder of the
  equilibration follows the same procedure as the dry system.

@@

  In order to fully match synthetic procedures, we created a crosslinking
  algorithm that can be applied to equilibrated structures. The purpose of 
  crosslinking
  is to maintain macroscopic alignment of the crystalline domains, 
  ensuring aligned, hexagonally packed pores. For that reason, we are not
  concerned with replicating the kinetics of the reaction, but instead emphasize
  the consistency of the final structure with experimental structural data. The
  algorithm was developed based on the known reaction mechanism. Crosslinking of
  this system is a free radical polymerization (FRP) taking place between
  terminal vinyl groups present on each of the three monomer tails. FRPs require
  an initiator which bonds to the system, meaning new atoms are introduced into
  the system. For simplicity, the initiator was simulated as hydrogen and made
  present in the simulation by including them in all possible locations where an 
  addition could occur as dummy atoms. The crosslinking procedure is carried out
  iteratively. During each iteration, bonding carbon atoms are chosen based on
  a distance cut-off. The topology is updated with new bonds and dummy hydrogen
  atoms are changed to appropriate hydrogen types. Head-to-tail addition was the
  only propagation mode considered due to its dominance in the real system.
  Direction of attack was not considered because the resultant mixture is
  racemic.

>>>>

%MRS3: this is a little disconnected to the contents as well.  You want
%the paper to explain the why of doing this research.  Paragraph gives background, but doesn't motivate.

@@

  %BJC : I don't know that this paragraph fits.  
  %MRS3: not really. Maybe take a look at Osuji's recent review (2016, I think) that addresses applications of nanostructured membranes and uses.  Could rewrite the introduction to be more about the general possibilities of nanostructured membranes, not just desalination.  It should be an introduction to what you will talk about in the paper, not to everything related.
  Improved RO membranes can make desalination more sustainable. Only 0.5 \% of
  the world's water is fresh. Desalination is an important source of potable
  water, necessary in order to keep up with demand generated by a rapidly growing
  global population. Although RO is currently the cheapest form of desalination,
  the cost must be further decreased in order to meet demand. A typical seawater
  desalination process requires feed pressures in the range of 800 to 1000 psi
  \cite{fritzmann_state---art_2007}. By designing better membranes for
  desalination, we can achieve higher water fluxes and reduce feed pressure
  requirements which will decrease the cost of fresh water production.
  \cite{humplik_nanostructured_2011}.

@@

%MRS3: I think that this paragraph probably should be couched in terms of what they COULD do, since are not widely used yet.
%MRS3: if talking about existing membranes, should cite performance.
  Nanostructured membranes can bypass many of the performance issues which
  plague traditional NF and RO membranes. One can accomplish targeted separations
  with high selectivity by tuning shape, size and functionality of the molecular
  building blocks which form these materials. As a result, solute rejecting pores
  can have their sizes tuned uniformly, resulting in strict size cut-offs.
  Entirely different mechanisms may govern transport in a given nanostructured
  material which can inspire novel separation techniques.

@@

  Self assembling lyotropic liquid crystals (LLCs) are a suitable candidate for
  aqueous separation applications. LLCs share the characteristic ability of
  nanostructured membrane materials to create highly ordered structures with the
  added benefits of low cost and synthetic techniques feasible for large scale
  production.  \cite{feng_scalable_2014} LLC systems created by the monomer 
  Na-GA3C11 (Fig.~\ref{fig:monomer}) have been extensively studied experimentally.
  \cite{smith_ordered_1997,zhou_supported_2005,resel_h2-phase_2000,feng_scalable_2014,feng_thin_2016}
  Neat liquid crystal monomer forms the thermotropic, Col\textsubscript{h}
  phase (Fig.\ref{fig:assembly}). The presence of ca. 10 wt \% water results in the
  H\textsubscript{II} phase. In both cases, monomers assemble into mesophases
  made of hexagonally packed, uniform size, cylinders with hydrophilic head
  groups oriented inward towards the cylinder center.  The hydrophilic region can
  act as a pore for aqueous separations.  One can envision tailoring the pore
  region for specific separations by changing the monomer chemistry.

@@

  %BJC: TODO: Scalable vector graphic for monomer structure, make bigger. Convert .svg to .eps
  %           Thicken slabs
  %           Make pore region on slab (blue) bigger
  %           Bump up atom sizes in atomistic render
  \begin{figure}
	\centering
	\begin{subfigure}{.3\textwidth}
		\centering
		\includegraphics[width=\textwidth]{NaGA3C11.png}
		\caption{}~\label{fig:monomer}
	\end{subfigure}
	\begin{subfigure}{.3\textwidth}
		\centering
		\includegraphics[width=\textwidth]{monomer_twocolor.png}
		\caption{}~\label{fig:atomistic_monomer}
	\end{subfigure}
	\begin{subfigure}{0.3\linewidth}
		\centering
		\includegraphics[width=\textwidth]{wedge.png}
		\caption{}~\label{fig:wedge}
	\end{subfigure}
		\begin{subfigure}{0.4\linewidth}
		\centering
		\includegraphics[width=\textwidth]{layer_blender.png}
		\caption{}~\label{fig:wedge_layer}
	\end{subfigure}
	\begin{subfigure}{0.4\linewidth}
		\centering
		\includegraphics[width=\textwidth]{hexagonal_packing.png}
		\caption{}~\label{fig:hex_packing_simple}
	\end{subfigure}
	\caption{The liquid crystal monomer Na-GA3C11 (a) rendered atomistically (b)
	exhibits wedge-like character (c). Monomer wedges assemble into disks (d) with
	hydrophilic head groups (blue) facing towards the disk center. The disks
	assemble into hexagonally packed columnar mesophases (e)}~\label{fig:assembly}
  \end{figure}

  Our current understanding of LLC systems is not rich enough to be able to
  precisely design membranes for specific separations. Over the past 20 years,
  H\textsubscript{II} phase LLC membrane studies have been limited primarily to
  Na-GA3C11 with some characterization done after minor structural modifications.
  Resel et al. varied the length of the monomer tails and the counterion used and
  observed its affect on pore spacing \cite{resel_structural_2000}.  Their study
  offers no insight into dynamics within the pore. In a later study of rejection
  performance, it was shown that membranes formed by Na-GA3C11 
%can not 
  cannot perform
  separations of solutes less than 1.2 nm in diameter because the pores are too
  large \cite{zhou_supported_2005}.  We do not yet understand how to controllably
  reduce the effective pore size or how to tune the chemical environment in the
  nanopores for effective water desalination and small organic molecule
  separations. The only source of predictive modeling for LLC systems have been
  macroscopic models which likely do not adequately describe transport at these
  length scales \cite{hatakeyama_water_2011}. It will be challenging to
  efficiently narrow down the large design space in a laboratory setting without
  a robust model.

  A molecular level understanding of LLC membrane structure, enabled by
  molecular dynamics simulations, 
%MRS3: suggest can provide, since we think it likely but cannot prove.
%will 
  can provide guidelines to reduce the large
  chemical space available to design monomers for creation of separation-specific
  membranes. A good molecular model should incorporate a detailed picture of the
  nanoscopic pore structure which will be crucial to understanding the role of
  monomer structure in solute transport and membrane design. Models resulting
  from molecular dynamics simulations will provide the required level of detail (Fig. \ref{fig:detail}).
  We can directly observe solute transport and suggest governing mechanisms. We
  can observe how the choice of head group may influence pore size for size
  exclusion driven separations. We can interchange counterions which may
  influence both the pore size and the strength of the Donnan potential which
  affects the degree to which the membrane can exclude charged species.

  % BJC: TODO: Make new rendering of wedges
  %            Change color scheme 
  \begin{figure}
  \centering
	\begin{subfigure}{0.45\linewidth}
		\centering
		\includegraphics[width=\textwidth]{nanopore_undetailed.jpg}
		\caption{}~\label{fig:undetailed_pore}
	\end{subfigure}
	\begin{subfigure}{0.45\linewidth}
		\centering
		\includegraphics[width=\textwidth]{detailed_pore.png}
		\caption{}~\label{fig:detailed_pore}
	\end{subfigure}
%MRS3: tweaked
%  \caption{(a) Our previous understanding of the pore structure allows us to speculate
%	   about separation behavior. (b) A detailed molecular model will allow us to
%	   directly observe solute transport. Here, four stacked layers of 5 monomers
%           are pictured atomistically. The hydrophilic region is in red and the 
%	   hydrophobic region is colored blue.}~\label{fig:detail}
  \caption{Our previous level of understanding of the pore structure (a) only allows us to speculate
	   about transport behavior in the pore. A detailed molecular model (b) will allow us to
	   directly observe solute transport. Here, four stacked layers of 5 monomers
           are pictured from atomistic simulations. The hydrophilic region is in red and the 
	   hydrophobic region is colored blue.}~\label{fig:detail}
  \end{figure}

@@

  The $\pi$-$\pi$ stacking interactions may be a driving force of self assembly
  in this system \cite{gazit_possible_2002}. Gas phase ab initio studies of
  benzene dimers have shown a clear energetic advantage for parallel displaced
  and T-shaped $\pi$-$\pi$ stacking conformations versus a sandwiched
  conformation ~\cite{sinnokrot_estimates_2002}. Substituted benzene rings
  exhibit an even stronger $\pi$-$\pi$ stacking attraction which favors the
  parallel displaced configuration in all cases except where the substitutions
  are extremely electron withdrawing.
%MRS3: perhaps ``in this study, we compare'', since the rest of the paper does not take place 
in the the future of the author. Apply throughout the paper.
  \cite{waller_hybrid_2006,ringer_effect_2006}. We will compare simulated X-ray
  diffraction patterns to experiment in order to deduce which stacking 
  configurations is most likely.

@@

  There remains the possibility that there is more than one metastable state
  associated with a given LLC system. Simulating a membrane atomistically will
%MRS3: rephrase.
%  require many atoms 
requires large systems
which limits the timescales acessible with MD. It is
  reasonable to expect that we will generate configurations which are kinetically
  trapped in a metastable free energy basin. We must be able to identify which
  state is produced experimentally.

@@

  While the Col\textsubscript{h} phase is described as dry, it has been
  suggested by experimentalists, in unpublished communications, that it is likely
  that small amounts of ambient water are leached into neat monomer.
  Experimentally, achieving a hexagonal phase with a completely dry system has
  proven difficult. If neat monomer is allowed to sit in ambient conditions, its
  color turns from transparent to slightly opaque and a hexagonal phase forms.
  Although we will not explore whether water is necessary for self assembly, we
  hypothesize that the hydrogen bonding network formed by the water may play a
  role in structuring the pores and holding together the hexagonal phase. We can
  use simulated X-ray diffraction patterns to see if there is any meaningful
  structural difference between a ``dry'' and ``wet'' system.

@@

  %BJC: Would it be better to describe each reflection as a numbered list (like with the questions above?)
  %MRS2: I like names better.
  We used experimental small-angle X-ray scattering (SAXS) data from
  %MRS3: note that ~ is a space, so you don't need to do Fig.~ ref.  The ~ keeps the Fig.~\ref together instead of being split over lines.
  \cite{feng_thin_2016} (Fig.~\ref{fig:SAXS}) and wide angle X-ray scattering
  (WAXS) data (Fig.~\ref{fig:WAXS}, produced as described in
  \cite{feng_scalable_2014}) for comparison to our model. We rely primarily on the 2D WAXS data
  since it encodes all structural details down to the sub-nm scale.  There are
  five major features of interest present in the 2D experimental pattern shown in
  Figure \ref{fig:WAXS}. 
%MRS3: break this into an enumerated list, rather than a wall of text.
The first is located at $q_z$ = 1.7 \angstrom$^{-1}$,
  corresponding to a real space separation of 3.7 \angstrom~.  The reflection is
  attributed to $\pi$-$\pi$ stacking between aromatic rings in the direction
  perpendicular to the membrane plane, or z-axis \cite{feng_scalable_2014}. For
  simplicity, this reflection will be referred to as R-$\pi$. A weak intensity
  line is located at exactly half the $q_z$ value of R-$\pi$ ($q_z$ = 0.85
  \angstrom$^{-1}$), corresponding to a real space periodic spacing of 7.4
  \angstrom~. This reflection has been interpreted as 2\textsubscript{1} helical
  ordering of aromatic rings along the z axis, meaning that if the positions of the
  aromatic rings can be traced by a helix, then for each full turn in the helix, one
  will encounter two aromatic rings. For this reason it will be referred to as
  R-helix. A third major reflection is marked by a low intensity ring located at
  r = 1.4 \angstrom$^{-1}$. The real space separation corresponds to 4.5 \angstrom~
  which is characteristic of the average spacing between packed alkane chains.
  This reflection will be called R-alkanes. Within R-alkanes, are four spots of
  higher relative intensity which will be called R-spots. All are located
  $\sim 37$ degrees from the $q_z$ axis in their respective quadrants. In many
  liquid crystal systems this can be explained by the tilt angle of the alkane
  chains with respect to the membrane plane. The final feature corresponds to the
  spacing and symmetry of the d\textsubscript{100} plane which can be related to
  the distance between pores.  The feature, which will be called R-pores, is
  characterized by dots along $q_z$ = 0. The spacing between dots is indicative
  of the hexagonal symmetry of the packed pores. The same information at higher
  resolution is obtained using a SAXS setup.

  %MRS3: this paragraph seems to be a little out of place: should go before a lot of the introductory material, since it lays out what you will be doing in the paper.

@@

  %BJC: TODO: update colorbar
  \begin{figure}
        \centering
        \begin{subfigure}[t]{0.43\linewidth}
                \centering
                \includegraphics[width=\linewidth]{SAXS.png}
                \caption{}\label{fig:SAXS}
        \end{subfigure}
        \begin{subfigure}[t]{0.47\linewidth}
                \centering
                \includegraphics[width=\linewidth]{WAXS_annotated.png} 
                \caption{}\label{fig:WAXS}
        \end{subfigure}
%MRS; would be good to have all feature labels agree: either can edit to put the names on, or have a key relating numbers to the names you use.
	\caption{(a) The repeat spacing in the 1D small angle X-ray scattering pattern
		is characteristic of hexagonal packing. The leading peak represents the
		distance between the d\textsubscript{100} planes. Using this distance, we know
		that the distance between pore centers is 4.12 nm. (b) 2D wide angle X-ray
		scattering gives details about repeating features less than 1 nanometer
		apart. Experimentalists have justified each of the 5 major reflections
		present as follows: (I) Aromatic head group $\pi-\pi$ stack 3.7 \AA~apart.
		(II) Monomers arrange vertically in a 2\textsubscript{1} helix. (III) Alkane
 		chain tails pack 4.5 \AA~apart. (IV) Monomer tails tilt with respect to the 
 		membrane plane. (V) As derived from SAXS, pores pack hexagonally and are 
		spaced 4.12 nm apart}
	\label{fig:SWAXS}
 \end{figure}

@@

  An ensemble of characteristic, low-energy vacuum monomer configurations
  were constructed by applying a simulated annealing process to a parameterized
  monomer. Monomers were cooled from 1000K to 50K over 10 nanoseconds.  A low
%MRS3: need to describe this in more detail, but can be in the supplementary. 
  energy configuration was randomly pulled from the trajectory and charges were
  reassigned using molcharge.
%MRS3: need to say what molcharge is.
  Using the new charges, the monomer system was
  annealed again and a random monomer configuration was pulled from the
  trajectory to be used for full system construction (Figure~\ref{fig:python}a).

@@

  A typical simulation volume contains four pores in a monoclinic unit cell,
  the smallest unit cell that maintains hexagonal symmetry when extended
  periodically. Each pore is made of twenty stacked monomer layers with periodic
  continuity in the z direction, avoiding any edge effects and creating an
  infinite length pore ideal for studying transport. A small number of layers is
  preferred in order to reduce computational cost and to allow us to look at
  longer timescales. Ultimately, we chose to build a system with 20 monomer
  layers in each pore in order to obtain sufficient resolution when simulating
  X-ray diffraction patterns. %This point will be explained in more detail later.
%MRS3: isn't there some observations you need at least 10?  
\subsection{Monomer Placement}

@@

  We chose the layer spacing for the initial configuration based on R-$\pi$ and
  then allowed the system to readjust during equilibration. Each monomer was rotated
  so the plane of the aromatic head groups would be coplanar with the xy plane.
  We explore two different initial layer spacings. The first is exactly equal to
  R-$\pi$ with layers placed so aromatic rings are stacked 3.7 \angstrom~ apart
  in the z-direction. A second system is explored with an initial layer spacing
  of 5 \angstrom. A third system with an initial layer spacing of 10 \angstrom was
  briefly explored. When layers are spaced out too far, they will collapse
  on each other while simultaneously slipping in between layers of adjacent pores
  which leads to an artificially thick membrane with pores spaced closely
  together. 
  Further details of simulations with large layer spacing are in the supplemental information.
%There will be no further discussion of this system but the interested
%  reader can learn more about it in the supplemental information.

  The relative interlayer orientation was chosen based on clues from
  diffraction data as well as the various known stacking modes of benzene and
  substituted benzene rings: sandwiched, parallel-displaced and T-shaped
  ~\cite{sinnokrot_estimates_2002} (\Cref{fig:sandwiched,fig:pd,fig:tshaped}).
  The T-shaped configuration was ruled out because its $\approx$ 5 \angstrom~
  equilibrium stacking distance ~\cite{sinnokrot_estimates_2002} is inconsistent
  with R-$\pi$. It is also infeasible for the monomers to orient in the T-shaped
  conformation because of the bulky tail groups. The system's preference towards
  the sandwiched vs. parallel displaced stacking modes will be explored in some
  detail.  Both have reported stacking distances near the R-$\pi$ value of 3.7
  \angstrom. Headgroups in our sandwiched initial configuration are stacked
  directly on top of each other while stacked headgroups in the parallel
  displaced initial configuration are offset by $180/nmon$ degrees where $nmon$
  equals the number of monomers per layer.

@@

  We developed equilibration schemes to create dry and wet configurations. Both
  schemes start with an initial configuration generated according to the previous
  guidelines. To create a dry configuration, we fix monomer head groups in the
  sandwiched or parallel-displaced configuration using position restraints with a
  force constant of 10$^6$ kJ mol$^{-1}$ nm$^{-2}$. We run a 50 ps simulation in the
  NVT ensemble which allows the monomer tails to settle without disrupting the
  ordering of the head groups. Doing so also mitigates system dependence on
  initial monomer configuration. Every 50 ps, we reduce the force constants by
  the square root of its previous value. Once the force constant is below 10 KJ
  mol$^{-1}$ nm$^{-2}$, the restraints are released linearly 
%MRS3: be more precise? you could reduce it from 10 in steps of 0.000001 kJ/mol, and that would be linear.
%until there is no more
%  restraining potential. 
until the restraining potential is zero.
The resulting unrestrained structure is allowed to
  equilibrate for 5 ns in the NPT ensemble with pressure controlled by the
  berendsen barostat. Next, we run long NPT equilibration simulations for at
  least 400 ns using the Parrinello-Rahman barostat with a time constant of 10 ps.

  In order to create a wet system, we solvated an initial configuration with
  water using \texttt{gmx solvate}. All water molecules placed outside the pore region are
  removed. Waters inside the pore region are randomly removed until the desired
  concentration of water in the pores is reached. The remainder of the
  equilibration follows the same procedure as the dry system.

@@

  In order to fully match synthetic procedures, we created a crosslinking
  algorithm that can be applied to equilibrated structures. The purpose of 
  crosslinking
  is to maintain macroscopic alignment of the crystalline domains, 
  ensuring aligned, hexagonally packed pores. For that reason, we are not
  concerned with replicating the kinetics of the reaction, but instead emphasize
  the consistency of the final structure with experimental structural data. The
  algorithm was developed based on the known reaction mechanism. Crosslinking of
  this system is a free radical polymerization (FRP) taking place between
  terminal vinyl groups present on each of the three monomer tails. FRPs require
  an initiator which bonds to the system, meaning new atoms are introduced into
  the system. For simplicity, the initiator was simulated as hydrogen and made
  present in the simulation by including them in all possible locations where an 
  addition could occur as dummy atoms. The crosslinking procedure is carried out
  iteratively. During each iteration, bonding carbon atoms are chosen based on
  a distance cut-off. The topology is updated with new bonds and dummy hydrogen
  atoms are changed to appropriate hydrogen types. Head-to-tail addition was the
  only propagation mode considered due to its dominance in the real system.
%MRS3: citation why it dominates the real system. 
  Direction of attack was not considered because the resultant mixture is
  racemic.

  Our implementation requires long simulation times 
%MRS3: specify long.  2 days?  2 months?
to achieve high degrees
  of crosslinking. For that reason we did not crosslink all systems tested, but only the most
  promising structure. We show that crosslinking does not significantly 
  change any of our drawn conclusions in Section 3.6.
